---
title: "Step 2: Optimal Embedding Dimension"   
output: html_document
---

## Optimal Embedding Dimension

Normalize the variables before performing EDM analyses.
Only use the six variables: incidence, temp, rain, rh, svpd, iobw
```{r}
all_vars <- c("last_yr_ir","ir","t2m","tp","rh","svpd","iobw")
all_vars_name <- c( "Last year's incidence","Incidence","Temperature", "Rain","RH", "SVPD","IOBW" )

climate_vars <- all_vars[-(1:2)]
local_climate_vars <- climate_vars[c(-5)]

# Normalize the whole dataset
data_country.norm <- scale(data_country[,all_vars], center = TRUE, scale = TRUE)

# Convert back to dataframe
data_country.norm <- as.data.frame(data_country.norm)

# Add the date back
time_vars <- c("time_id","date","year","month","country")
data_country.norm <- cbind(data_country[,time_vars], data_country.norm)

# library size
lib_ccm <- c(1,NROW(data_country.norm))
```

### Simplex projection

We will use simplex projection to determine the dimensionality of the variables and if the variables are predictable (i.e., if governed by determinisitc processes). 

Option 1: We use EmbedDimension() to determine the optimal embedding dimension, E.
```{r}
E.opt <- lapply(climate_vars, function(var) {
  EmbedDimension( dataFrame = data_country.norm[, c("time_id", all_vars)],    # input data
                  lib       = lib_ccm, # portion of data to train
                  pred      = lib_ccm, # portion of data to predict
                  maxE      = 4,
                  columns   = var,
                  target    = "ir" )
})
names(E.opt) <- climate_vars
opt_E <- sapply(E.opt, function(df) {
  df$E[which.max(unlist(df$rho))]
})
opt_E
```
### Evidence of Nonlinear Dynamics
We use PredictNonlinear() to test nonlinear dynamics.
```{r}

find_max_Theta <- function(df) {
  integer_thetas <- df$Theta[df$Theta == as.integer(df$Theta)]
  if (length(integer_thetas) > 0) {
    max_rho_index <- which.max(df$rho[df$Theta %in% integer_thetas])
    return(df$Theta[df$Theta %in% integer_thetas][max_rho_index])
  } else {
    return(NA) 
  }
}

theta.rho  <- lapply(climate_vars, function(var) {
  PredictNonlinear( dataFrame = data_country.norm[, c("time_id", all_vars)],    # input data
                    lib       = lib_ccm, # portion of data to train
                    pred      = lib_ccm, # portion of data to predict, 
                    E         = opt_E[var],
                    columns   = var,
                    target    = "ir" )
})
names(theta.rho) <- climate_vars

max_theta.rho <- sapply(theta.rho, find_max_Theta)
print(max_theta.rho)

# Plot Nonlinear
par(mfrow = c(2, 3), mar = c(4,4,4,4), mgp = c(1,0.5,0), oma = c(0,0,0,0))

for (var in names(theta.rho)) {
  plot(theta.rho[[var]]$Theta, theta.rho[[var]]$rho, plot.main=NULL, type = "l",lwd=2,
       axes = FALSE,
       xlab = "", ylab = "")
  box(col="black")
  axis(1, col="black", col.ticks="black", col.axis="black", cex.axis=1, tck = 0.02)
  mtext("Nonlinearity (theta)", side=1, line=2, col="black", cex=0.8)
  
  axis(2, col="black", col.ticks="black", col.axis="black", cex.axis=1, tck = 0.02)
  mtext("Forecast Skill (rho)", side=2, line=2, col="black", cex=0.8)
  abline(v = max_theta.rho[var], col="black", lwd=3, lty=2)
  title(var,cex.main=2)
}

output_name <- paste("02_output/01_parameter_selection/03_Nonlinear/", country_index, "_", country_name, sep="")
file_name = paste(output_name, ".pdf", sep="")
dev.copy(pdf, file=file_name, width = 8, height = 5)
dev.off()
```
### Forecast interval accuracy
We use PredictInterval() to test nonlinear dynamics.
```{r}
Tp.rho  <- lapply(climate_vars, function(var) {
  PredictInterval( dataFrame  = data_country.norm[, c("time_id", all_vars)],    # input data
                    lib       = lib_ccm, # portion of data to train
                    pred      = lib_ccm, # portion of data to predict, 
                    E         = opt_E[var],
                    columns   = var,
                    target    = "ir")
})
names(Tp.rho) <- climate_vars

```

Here, we can save our progress.
```{r}
# save E and theta
opt_E <- t(data.frame(opt_E))
colnames(opt_E) <- paste0("E_", colnames(opt_E))

max_theta.rho <- t(data.frame(max_theta.rho))
colnames(max_theta.rho) <- paste0("theta_", colnames(max_theta.rho))

ccm_parameter <- cbind(ccm_parameter_lag, opt_E, max_theta.rho)
ccm_parameter <- as.data.frame(ccm_parameter)

ccm_parameter$country <- country_name
rownames(ccm_parameter) <- 1

ccm_parameter_all <- rbind(ccm_parameter_all,ccm_parameter)
# rename row
rownames(ccm_parameter_all) <- 1:nrow(ccm_parameter_all)

```