---
title: "Figure 3ab"   
---

##  ----------------------------Convergent Cross-Mapping (Figure 3a)

If necessary, we can clear the environment to start with a clean workspace.
```{r}
rm(list=ls())
```

Help functions.
```{r}
library(ggplot2)
```

Here, we can load previous results.
```{r}
ds_all   <- read.csv("data/02_uEDM_ir.csv", header = T)
climates <- unique(ds_all$climate)
countrys <- unique(ds_all$country)
models <- c("Seasonal","Ebisuzaki")
```

### State Space Reconstruction

```{r Set general parameters}
t2m_color <- "red"
tp_color <- "skyblue2"
rh_color <- "green3"
svpd_color <- "goldenrod2"
ir_color <- "black"
season <- c(t2m_color,tp_color,rh_color,svpd_color)
plot_box <- c("p1","p2","p3","p4")
```

### Data

Add lagged and averaged lagged variables.
```{r}
plot_list = list() 

for (j in 1:2){
  for (i in 1:4){
    ds <- ds_all %>%
      filter(climate==climates[i]) %>%
      filter(model==models[j]) %>%
      mutate(country = factor(country, levels =  rev(countrys)))
    
    p <- ggplot(ds) + 
      geom_boxplot(aes(x=ccm_null, y=country), color = season[i],size = 1) +
      geom_point(aes(x=ccm_driver, y=country, color = ir_color), shape = 1, color = "black",size = 4, na.rm = T) +
      coord_cartesian(xlim = c(0,1)) +
      scale_x_continuous(breaks=c(0,0.5,1)) +
      labs(x = " ", y=" ", title = climates[i])+
      theme_test(base_size = 15)+ 
      theme(legend.position = "none",
            axis.text.y = element_blank(),
            axis.ticks.y= element_blank()) + 
      theme(axis.text = element_text(color = 'black'),
            plot.margin = margin(1,0.5,0.5,2.5,'cm'),
            panel.border = element_rect(size = 1),
            axis.title = element_text(face = 'bold'),
            plot.title = element_text(face = 'bold',size=13,hjust = 0.5))
    num <- (j-1)*4+i
    plot_list[[num]] <- p 
    
}
}
fg1_1 <- plot_list[[1]] %>% ggplotGrob()
fg1_2 <- plot_list[[2]] %>% ggplotGrob()
fg1_3 <- plot_list[[3]] %>% ggplotGrob()
fg1_4 <- plot_list[[4]] %>% ggplotGrob()
cg1 <- cowplot::plot_grid(fg1_1, fg1_2, fg1_3, fg1_4, nrow =1)

fg2_1 <- plot_list[[5]] %>% ggplotGrob()
fg2_2 <- plot_list[[6]] %>% ggplotGrob()
fg2_3 <- plot_list[[7]] %>% ggplotGrob()
fg2_4 <- plot_list[[8]] %>% ggplotGrob()
cg2 <- cowplot::plot_grid(fg2_1, fg2_2, fg2_3, fg2_4, nrow =1)

cairo_pdf(file = "output/fig_3a.pdf", width = 10, height = 10)
cowplot::plot_grid(cg1, cg2, ncol = 1)
dev.off()
```

## Paired tests for independence of two sets of data
```{r}
test_all   <- read.csv("data/03_uEDM_ir_T_test.csv", header = T)

# test for normality
#p<0.05, data not normal
library(reshape2)
#tapply(ds_test$rho, ds_test$group, shapiro.test)

ds_test <- test_all %>%
      filter(climate==climates[1]) %>%
      filter(group!=2) 

wilcox.test(ds_test$rho,ds_test$group,alternative="two.sided")
#reslut$p.value


```

### Plotting multivariate EDM forecast improvements.
### When the EDM model includes driver variables, a causal effect is demonstrated if the EDM prediction skill (ρ) improves. Quantified by Δρ = ρ (with driver) - ρ (no driver, ir_t,ir_t-2,ir_t-4 only), where ρ is the Pearson correlation between observations and EDM predictions.
```{r}
delta_all   <- read.csv("data/04_mEDM_delta_ir-t.csv", header = T)
cols_all <- unique(delta_all$cols)
delta <- delta_all %>%
  mutate(cols = factor(cols, levels =  cols_all))

ggplot(delta) + 
  geom_boxplot(aes(x=cols, y=rho), color = "skyblue3", size = 0.7) +
  #geom_boxplot(aes(x=cols, y=rmse), color = "indianred3", size = 0.7) +
  geom_hline(yintercept = 0,linetype="dotted", color = 'gray40',size = 1) + 
  labs(x = "Driver in EDM", y="Cross map skill (ρ)")+
  theme_test(base_size = 15)+ 
  #theme(legend.position = "none") + 
  theme(axis.text = element_text(color = 'black'),
        axis.text.x = element_text(angle = 35,vjust = 0.7,hjust = 0.5),
        #panel.border = element_rect(size = 1),
        axis.title = element_text(face = 'bold'),
        plot.title = element_text(face = 'bold',size=17,hjust = 0.5))

#dev.copy(pdf, paste("output/fig_3b.pdf", sep=""), width = 8, height = 4)
#dev.off()

```

## smap mEDM
```{r}
delta_all   <- read.csv("data/05_mEDM_delta_ir_lag.csv", header = T)
embedding_all <- unique(delta_all$embedding)
delta <- delta_all %>%
  mutate(embedding = factor(embedding, levels =  embedding_all))

max_rho <- delta %>%
  filter(embedding=="t2m+tp+rh") %>%
  filter(type=="ir_yr")

m_rho <- median(max_rho$rho)
m_rmse <- median(max_rho$rmse)

p1 <- ggplot(delta,aes(x=embedding, y=rho)) + 
  stat_boxplot(aes(fill=type),geom="errorbar",width=0.1,size=0.6,position=position_dodge(0.6))+
  geom_boxplot(aes(fill=type), position=position_dodge(0.6), size=0.5,               width=0.5)+
  scale_fill_manual(values = c("skyblue3","indianred3")) +
  geom_hline(yintercept = m_rho,linetype="dotted", color = 'gray40',size = 1) + 
  labs(x = "Driver in EDM", y="Cross map skill (ρ)")+
  theme_test(base_size = 15)+ 
  theme(axis.text = element_text(color = 'black'),
        axis.text.x = element_text(angle = 35,vjust = 0.7,hjust = 0.5),
        axis.title = element_text(face = 'bold'),
        plot.title = element_text(face = 'bold',size=17,hjust = 0.5))

p2 <- ggplot(delta,aes(x=embedding, y=rmse)) + 
  stat_boxplot(aes(fill=type),geom="errorbar",width=0.1,size=0.6,position=position_dodge(0.6))+
  geom_boxplot(aes(fill=type), position=position_dodge(0.6), size=0.5,               width=0.5)+
  scale_fill_manual(values = c("skyblue3","indianred3")) +
  geom_hline(yintercept = m_rmse,linetype="dotted", color = 'gray40',size = 1) + 
  labs(x = "Driver in EDM", y=" Root Mean Square Error ")+
  theme_test(base_size = 15)+ 
  theme(axis.text = element_text(color = 'black'),
        axis.text.x = element_text(angle = 35,vjust = 0.7,hjust = 0.5),
        axis.title = element_text(face = 'bold'),
        plot.title = element_text(face = 'bold',size=17,hjust = 0.5))

fg1 <- p1 %>% ggplotGrob()
fg2 <- p2 %>% ggplotGrob()

cairo_pdf(file = "output/fig_3b.pdf", width =15, height = 4)
cowplot::plot_grid(fg1, fg2, nrow = 1)
dev.off()

```